---
title: "Analysis Sample and Effect Size 11"
author: Simon Stephan
output: html_document
---

```{r setup, include=FALSE}
# code that automatically sets the WD to source file location
#rstudioapi::getActiveDocumentContext
#setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
getwd() # check wd

# packages
library(ez)
library(reshape2)
library(reshape)
library(ggplot2)
library(plyr)
library(pastecs)
library(data.table)
library(psych)

library(showtext)

font_add_google("Poppins", "Poppins")
font_add_google("Roboto Mono", "Roboto Mono")
showtext_auto()
```

```{r}
# read data 
#tdata.wide <- read.delim("dummy_data.txt", header=TRUE, sep="\t", na.strings="NA", dec=".", strip.white=TRUE)
tdata <- read.delim("Exp3_data.txt", header=TRUE, sep="\t", na.strings="NA", dec=".", strip.white=TRUE)
```

# Demographics
```{r}
mean(tdata$Age)
sd(tdata$Age)

colnames(tdata)[12] <- "Sex" # rename column listing subjects' sex
table(tdata$Sex) # 1 = male, 2 = female, 3 = non-binary
```

# Data preparation
```{r, message=FALSE}
# declare factor

tdata$High_Strength_Component <- factor(tdata$High_Strength_Component, levels = c("Terminal", "Root"))
tdata$Observation_Order <- factor(tdata$Observation_Order, levels = c("unobserved_first", "observed_first"))

tdata$diag_prob_rating <- (tdata$diag_prob_rating - 1) * 0.1

tdata_attributions <- subset(tdata, select = c(1:6))
tdata_diagnostics <- subset(tdata, select = c(1:4,7))
tdata_diagnostics$High_Strength_Component <- factor(tdata_diagnostics$High_Strength_Component, levels = c("Root", "Terminal"))

# rename columns 
colnames(tdata_attributions)
colnames(tdata_attributions)<- c("sID", "High_Strength_Component", "Target", "Observation_Order", "1_1_u_1", "1_u_u_1")

```




```{r, message=FALSE}
tdata_attributions_long <- melt(tdata_attributions, id=c("sID", "High_Strength_Component", "Target", "Observation_Order"))

tdata_attributions_long$variable <- factor(tdata_attributions_long$variable, levels = c("1_u_u_1", "1_1_u_1"))

# 5. recode dependent variables (because Subj. answered on a 1 to 11 scale but we want 0 to 10)
tdata_attributions_long$value <- (tdata_attributions_long$value - 1) * 0.1

tdata_attributions_long["title"] <- c(rep("Human judgments", length(tdata_attributions_long$sID)))
```

# Make bar charts 

## Chart showing the predictions 

```{r, echo = FALSE}
# create a theme to handle the design 
myTheme <- theme(plot.title = element_text(face="bold", size = 12),
        axis.title.x = element_text(face = "bold", size = 10),
        axis.title.y = element_text(face = "bold", size = 10),
        axis.text.x = element_text(size = 6, angle = 45, hjust = 1), 
        axis.text.y = element_text(size = 4, angle = 90),
        legend.text = element_text(size = 12),
        legend.title = element_text(face = "bold", size = 12),
        strip.text.x = element_text(size = 12),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line.x = element_line(colour = "black"), 
        axis.line.y = element_line(colour = "black"),
        axis.text = element_text(colour ="black"), 
        axis.ticks = element_line(colour ="black"))

```

## predictions based on integration of information from both potential causes


## predictions 

```{r}
# derive the model predictions for the condition in which the terminal path is strong

##### Causal strength parameters

# base rates of root causes C and A
bc <- 0.7
ba <- bc

# strength between root causes and mechanism nodes
wcmc <- 0.1
wama <- wcmc

# strength of mechanism component
wmce <- 0.9 # the strong path
wmae <- wmce


alpha_symmetric <- 0.5 # for x_u_1_u_x




##### The 2 different test cases we need to model

#############################################
# legend: 
# t = telegraph (superior mech component)
# p = pony rider (inferior mech component)
# 1 = variable present 
# 0 = variable absent 
# u = variable unobserved
#############################################

#     C_Mc_E_Ma_A

#  1) 1_u_1_u_1
#  2) 1_1t_1_u_1


##################################################

# function for generalized power model of singular causation 

generalized_model <- function(target_power, alternative_power, alpha) 
{
  # formula 
  (target_power - target_power * alternative_power * alpha)/(target_power + alternative_power - target_power * alternative_power)
}

# 1) 1_u_1_u_1
power_target <- wcmc * wmce
power_alt <- wama * wmae
alpha <- alpha_symmetric

singular_1 <- generalized_model(power_target, power_alt, alpha)


# 1) 1_1_1_u_1
power_target <- wmce # simply the determined link strength
power_alt <- wama * wmae
alpha <- alpha_symmetric


singular_2 <- generalized_model(power_target, power_alt, alpha)

singular_1
singular_2
```


```{r}
# derive the model predictions for the condition in which the root path is strong 

##### Causal strength parameters

# base rates of root causes C and A
bc <- 0.7
ba <- bc

# strength between root causes and mechanism nodes
wcmc <- 0.9
wama <- wcmc

# strength of mechanism component
wmce <- 0.1 # the strong path
wmae <- wmce


alpha_symmetric <- 0.5 # for x_u_1_u_x




##### The 2 different test cases we need to model

#############################################
# legend: 
# t = telegraph (superior mech component)
# p = pony rider (inferior mech component)
# 1 = variable present 
# 0 = variable absent 
# u = variable unobserved
#############################################

#     C_Mc_E_Ma_A

#  1) 1_u_1_u_1
#  2) 1_1t_1_u_1


##################################################

# function for generalized power model of singular causation 

generalized_model <- function(target_power, alternative_power, alpha) 
{
  # formula 
  (target_power - target_power * alternative_power * alpha)/(target_power + alternative_power - target_power * alternative_power)
}

# 1) 1_u_1_u_1
power_target <- wcmc * wmce
power_alt <- wama * wmae
alpha <- alpha_symmetric

singular_3 <- generalized_model(power_target, power_alt, alpha)


# 1) 1_1_1_u_1
power_target <- wmce # simply the determined link strength
power_alt <- wama * wmae
alpha <- alpha_symmetric


singular_4 <- generalized_model(power_target, power_alt, alpha)

singular_3
singular_4
```

```{r}
# create a data file containing the predictions 

High_Strength_Component <- c(rep("Root",4), rep("Terminal",4))
Observation_Order <- c(rep(c("unobserved_first", "unobserved_first", "observed_first", "observed_first"),4))

variable <- c(rep(c("1_u_u_1", "1_1_u_1", "1_1_u_1", "1_u_u_1"),2))

value <- c(singular_3, singular_4, singular_4, singular_3, singular_1, singular_2, singular_2, singular_1)

title <- c(rep("Ordinal predictions: integration of target and alternative cause", length(value)))

preds <- data.frame(High_Strength_Component, Observation_Order, variable, value, title)

preds$High_Strength_Component <- factor(preds$High_Strength_Component, levels = c("Terminal", "Root"))
preds$Observation_Order <- factor(preds$Observation_Order, levels = c("unobserved_first", "observed_first"))
preds$variable <- factor(preds$variable, levels = c("1_u_u_1", "1_1_u_1"))
```

```{r}
library(scales) 

theme_set(theme_bw(base_size = 12, base_family = "Poppins"))


ggplot(preds, aes(x=variable, y=value, group = High_Strength_Component, colour= High_Strength_Component)) + 
  ggtitle("a) Predictions") +
  coord_cartesian(ylim=c(0,1)) +
  #coord_flip()+
  #scale_y_continuous(limits = c(0,1), breaks=seq(0, 1, 0.25), expand = c(0,0)) +
  #scale_x_discrete(labels=c("Direct: one link", "Iindirect: two links", "Indirect: five links")) +
  #stat_summary(fun.y = mean, geom = "bar", position = "dodge", colour = "black") +
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0, position=position_dodge(0.1), size = 1) +
  stat_summary(aes(y = value), fun.y=mean, geom="line", position=position_dodge(0.1), size = 1, alpha = 0.7)+
  stat_summary(aes(y = value), fun.y=mean, colour="black", fill = c("#005083", "#005083", "#f0b64d", "#f0b64d"), geom="point", shape = 21, size = 3, position=position_dodge(0.1))+
  #geom_jitter(alpha = 0.2, colour = "black") +
  #stat_summary(fun.y = median, geom = "point", shape = 23, size = 3, position = position_dodge(width = 0.90), colour = "black") +
  #geom_hline(yintercept=0.9) + 
  #ggtitle("Means (95% bootstr. CIs)") +
    theme(
      legend.position = "none",
      axis.title = element_text(size = 13),
      axis.text.y = element_text(angle = 90, color = "black", size = 12, hjust = 0.2),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      panel.grid = element_blank(),
      #panel.grid = element_line(colour="grey", linetype="dashed")
      )+
  labs(x = "Observed singular case", y = "Confidence that C = 1 caused E = 1") +
  scale_color_manual(name  ="High-strength comp.", labels=c("terminal path", "root path"), values=c("#005083", "#f0b64d")) +
  #scale_colour_discrete(name  ="Diagnostic judgment", labels=c("correct", "incorrect")) +
  theme(legend.position="top")
  #theme(legend.position = c(0.2, 0.2),
   #     legend.title = element_text(size = 8),
    #    legend.text = element_text(size = 8),
     #   legend.background = element_rect(fill=alpha("grey", 0.4)))


#ggsave("predictions1.svg",width=5,height=4)
#ggsave("predictions1.pdf",width=5,height=4)


```

```{r}
# plot the interaction effect together with its 95% CI 

t <- qt(0.975, 142, lower.tail = TRUE, log.p = FALSE)
#t

effect <- "Delta D"
D_Mdiff <- round((singular_2 - singular_1) - (singular_4 - singular_3),1)
SE <- 0.00
CI <- SE*t
CI_low <- D_Mdiff - CI
CI_up <- D_Mdiff + CI

DMdiff <- data.frame(effect, D_Mdiff, CI_low, CI_up)

ggplot(DMdiff, aes(x=effect, y=D_Mdiff)) +
    #coord_cartesian(ylim=c(-0.02,0.15)) +
    #scale_y_continuous(limits = c(0,0.2), breaks=seq(0, 0.2, 0.1), expand = c(0,0)) +
    #geom_line() +
    geom_errorbar(width=.0, aes(ymin=DMdiff$CI_low, ymax=DMdiff$CI_up)) +
    geom_point(shape=17, size=3, fill="red", color = "black") +
    #ylim(-0.02,0.3)+
    scale_y_continuous(limits=c(-0.02, 0.8), breaks = seq(-0.02, 0.8, 0.2))+
    geom_hline(yintercept=0, linetype="dashed", color = "red")+
   theme(
      legend.position = "none",
      axis.line.y = element_line(colour = "black"),
      axis.title.x = element_blank(),
      axis.ticks.y = element_line(color = "black"),
      axis.ticks.x = element_blank(),
      #axis.ticks.y = element_line(colour = "black"),
      axis.text.x = element_text(size = 10, color = "black"),
      axis.title.y = element_blank(),
      axis.text.y = element_text(size = 12, color = "black"),
      panel.grid = element_blank(),
      panel.border = element_blank(),
      #panel.grid = element_line(colour="grey", linetype="dashed")
      )

#ggsave("interaction_main_pred.svg",width=1,height=2)
#ggsave("interaction_main_pred.pdf",width=1,height=2)

```


## Chart showing human diagnostic judgments 

```{r}

diag_judgments <- ggplot(tdata_diagnostics, aes(High_Strength_Component, diag_prob_rating)) + 
  #facet_grid(. ~ Parameter_Info) +
  coord_cartesian(ylim=c(0,1)) + 
  scale_y_continuous(limits = c(0,1), breaks=seq(0, 1, 0.2), expand = c(0,0)) +
  #scale_x_discrete(labels=c("Direct: one link", "Iindirect: two links", "Indirect: five links")) +
  stat_summary(fun.y = mean, geom = "bar", position = "dodge", colour = "black", fill = "#bcbddc") +
  #geom_errorbar(aes(ymin= dat$means - dat$conf, 
  #                  ymax=dat$means + dat$conf), position = position_dodge(width = 0.10), width=0.2)+
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", position = position_dodge(width = 0.90), width = 0.2) +
  #ggtitle("Means (95% bootstr. CIs)") +
  theme(plot.title = element_text(face="bold", size = 12),
        axis.title.x = element_text(face = "bold", size = 12),
        axis.title.y = element_text(face = "bold", size = 12),
        axis.text.x = element_text(face = "bold", size = 12), 
        axis.text.y = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(face = "bold", size = 12),
        strip.text.x = element_text(size = 12),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line.x = element_line(colour = "black"), 
        axis.line.y = element_line(colour = "black"),
        axis.text = element_text(colour ="black"), 
        axis.ticks = element_line(colour ="black")) + 
  labs(x = "Location of highest strength", y = "Probability of Mechanism Activation") +
  scale_fill_manual(values=c("#bcbddc", "#3182bd")) +
  theme(legend.position="none")

diag_judgments


#ggsave("diagnostic_judgments.png",width=5,height=6)
```

```{r}

tdata_diagnostics$High_Strength_Component <- factor(tdata_diagnostics$High_Strength_Component, levels = c("Terminal", "Root"))

ggplot(tdata_diagnostics, aes(High_Strength_Component, diag_prob_rating, color = High_Strength_Component, fill = High_Strength_Component)) + 
  ggtitle("c) Subjects' estimations of the probability that the alternative cause's mechanism was active")+
  #coord_cartesian(ylim=c(-0.1,2)) + 
  scale_y_continuous(limits = c(-0.05, 1.06), breaks=seq(0, 1, 0.25), expand = c(0,0)) +
  scale_x_discrete()+
  #scale_x_discrete(labels=c("terminal path", "root path")) +
  #stat_summary(fun.y = mean, geom = "bar", position = "dodge", colour = "black", fill = "#bcbddc") +
  #geom_errorbar(aes(ymin= dat$means - dat$conf, 
  #                  ymax=dat$means + dat$conf), position = position_dodge(width = 0.10), width=0.2)+
  #annotate("rect", xmin=1.7, xmax=2.3, ymin=0.50, ymax=1.05, fill="#31a354", alpha=0.6, color = "black") +
  #annotate("rect", xmin=1.7, xmax=2.3, ymin=0.0, ymax=0.50, fill="#de2d26", alpha=0.6, color = "black") +
  #annotate("rect", xmin=1.7, xmax=2.3, ymin=0.95, ymax=1.05, color="#31a354", fill = "white", size = 1.5) +
  #annotate("rect", xmin=1.7, xmax=2.3, ymin=0.45, ymax=0.55, color="#de2d26", fill = "white", size = 1.5) +
  geom_jitter(width = 0.3, height = 0.02, alpha = 0.6) +
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", position = position_dodge(width = 0.90), width = 0, size = 1) +
  stat_summary(fun.y=mean, colour="grey20", geom="point", shape = 21, size = 3)+
  #ggtitle("Means (95% bootstr. CIs)") +
  theme(
  legend.position = "none",
      axis.title = element_text(size = 13.5),
      axis.text.y = element_text(angle = 90, color = "black", size = 12.5, hjust = 0.2),
      axis.text.x = element_text(color = "black", size = 12.5),
      #axis.title.x = element_blank(),
      panel.grid = element_blank(),
      #panel.grid = element_line(colour="grey", linetype="dashed")
      )+
  #theme(axis.text.x = element_text(size = 10, angle = 0, hjust = 0.5))+
  labs(x = "High-strength component", y = "Estimated probability") +
  scale_color_manual(values=c("#005083", "#f0b64d")) +
  scale_fill_manual(values=c("#005083", "#f0b64d")) +
  theme(legend.position="none")


#ggsave("diagnostic_judgments_jit.pdf",width=5,height=4)
```

```{r}
by(tdata_diagnostics$diag_prob_rating, list(tdata_diagnostics$High_Strength_Component), stat.desc , basic = FALSE)

```


## Chart showing human singular causation judgments

how within-subject error bars are to be computed can be found here: http://www.cookbook-r.com/Graphs/Plotting_means_and_error_bars_(ggplot2)/ 


```{r}
library(scales) 

theme_set(theme_bw(base_size = 12, base_family = "Poppins"))

ggplot(tdata_attributions_long, aes(x=variable, y=value, group = High_Strength_Component, colour= High_Strength_Component)) + 
  ggtitle("b) Human singular causation judgments") +
  coord_cartesian(ylim=c(0,1)) +
  #coord_flip()+
  #scale_y_continuous(limits = c(0,1), breaks=seq(0, 1, 0.25), expand = c(0,0)) +
  #scale_x_discrete(labels=c("Direct: one link", "Iindirect: two links", "Indirect: five links")) +
  #stat_summary(fun.y = mean, geom = "bar", position = "dodge", colour = "black") +
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0, position=position_dodge(0.1), size = 1) +
  stat_summary(aes(y = value), fun.y=mean, geom="line", position=position_dodge(0.1), size = 1, alpha = 0.7)+
  stat_summary(aes(y = value), fun.y=mean, colour="grey20", fill = c("#005083", "#005083", "#f0b64d", "#f0b64d"), geom="point", shape = 21, size = 3, position=position_dodge(0.1))+
  #geom_jitter(alpha = 0.2, colour = "black") +
  #stat_summary(fun.y = median, geom = "point", shape = 23, size = 3, position = position_dodge(width = 0.90), colour = "black") +
  #geom_hline(yintercept=0.9) + 
  #ggtitle("Means (95% bootstr. CIs)") +
  theme(
      legend.position = "none",
      axis.title = element_text(size = 12),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      panel.grid = element_blank(),
      axis.title.y = element_text(color = "white"), 
      axis.text.y = element_text(color = "white")
      #panel.grid = element_line(colour="grey", linetype="dashed")
      )+
  labs(x = "Observed singular case", y = "Confidence that C = 1 caused E = 1") +
  scale_color_manual(name  ="High-strength comp.", labels=c("terminal path", "root path"), values=c("#005083", "#f0b64d")) +
  #scale_colour_discrete(name  ="Diagnostic judgment", labels=c("correct", "incorrect")) +
  theme(legend.position="top")
  #theme(legend.position = c(0.2, 0.2),
  #      legend.title = element_text(size = 8),
  #      legend.text = element_text(size = 8),
  #      legend.background = element_rect(fill=alpha("grey", 0.4)))


#ggsave("human_jugments.pdf",width=5,height=4)
```





# Statistical Analyses

## Descriptive Statistics
```{r}
#library(pastecs)
by(tdata_attributions_long$value, list(tdata_attributions_long$variable, tdata_attributions_long$High_Strength_Component), stat.desc , basic = FALSE)

#by(tdata2$DeltaP, list(tdata2$Conclusion), stat.desc , basic = FALSE)

#tdata2<-subset(tdata2, tdata2[,5]!="-Inf") 
#tdata2<-subset(tdata2, tdata2[,5]!="Inf") 
#tdata2<-subset(tdata2, tdata2[,5]!="NaN") 

#by(tdata2$Power, list(tdata2$Conclusion), stat.desc , basic = FALSE)


```

```{r}
# compute correlation among repeated measures in each group 

root <- subset(tdata_attributions, High_Strength_Component == "Root")
terminal <- subset(tdata_attributions, High_Strength_Component == "Terminal")

cor(root$`1_u_u_1`, root$`1_1_u_1`)

cor(terminal$`1_u_u_1`, terminal$`1_1_u_1`)



```

```{r}
library(afex)
library(emmeans)

a1 <- aov_car(value ~ High_Strength_Component*variable + Error(sID/(variable)), tdata_attributions_long)
a1
```


```{r}
# same ANOVA as before
lmeModel <- lmer(value ~ High_Strength_Component*variable + (1|sID), data=tdata_attributions_long)

# follow-up analysis 

ls1 <- lsmeans(a1, c("variable", "High_Strength_Component")) # joint evaluation (basically gives the same table)
ls1
```


```{r}
############### 
# a conditional analysis 

ls2 <- lsmeans(a1, c("variable"), by = "High_Strength_Component") # group means by between-condition
ls2

# simple main effects 
pairs(ls2) # compares rep-measure differences separately for each between-factor level

# interaction contrast 
pairs(pairs(ls2), by = NULL)

#test(pairs(pairs(ls2), by = NULL), joint = TRUE) # This reproduces the F-Value of the ANOVA interaction


#lsmip(a1, High_Strength_Component ~ variable) # lsemans can also produce graphs
```


```{r}
# plot the interaction effect together with its 95% CI 

t <- qt(0.975, 158, lower.tail = TRUE, log.p = FALSE)
#t

effect <- "Delta D"
D_Mdiff <- 0.115
SE <- 0.035
CI <- SE*t
CI_low <- D_Mdiff - CI
CI_up <- D_Mdiff + CI

DMdiff <- data.frame(effect, D_Mdiff, CI_low, CI_up)

ggplot(DMdiff, aes(x=effect, y=D_Mdiff)) +
    #coord_cartesian(ylim=c(-0.02,0.15)) +
    #scale_y_continuous(limits = c(0,0.2), breaks=seq(0, 0.2, 0.1), expand = c(0,0)) +
    #geom_line() +
    geom_errorbar(width=.0, aes(ymin=DMdiff$CI_low, ymax=DMdiff$CI_up)) +
    geom_point(shape=17, size=3, color = "black") +
    #ylim(-0.02,0.2)+
    scale_y_continuous(limits=c(-0.02, 0.8), breaks = seq(-0.02, 0.8, 0.2))+
    geom_hline(yintercept=0, linetype="dashed", color = "red")+
theme(
      legend.position = "none",
      axis.line.y = element_line(colour = "black"),
      axis.title.x = element_blank(),
      axis.ticks.y = element_line(color = "black"),
      axis.ticks.x = element_blank(),
      #axis.ticks.y = element_line(colour = "black"),
      axis.text.x = element_text(size = 10, color = "black"),
      axis.title.y = element_blank(),
      axis.text.y = element_text(size = 12, color = "black"),
      panel.grid = element_blank(),
      panel.border = element_blank(),
      #panel.grid = element_line(colour="grey", linetype="dashed")
      )

#ggsave("interaction.pdf",width=1,height=2)


```




```{r}
# correlation between P(MA = 1) ratings and singular causation ratings (should be negative)
rest <- subset(tdata, select = c(1:7))
colnames(rest)<- c("sID", "High_Strength_Component", "Target", 
                                   "Observation_Order", "1_1_u_1", "1_u_u_1", "diagnostic_judgment")

rest_long <- melt(rest, id=c("sID", "High_Strength_Component", 
                                                            "Target", "Observation_Order", "diagnostic_judgment"))

rest_long$value <- (rest_long$value - 1) * 0.1

rest_11u1 <- subset(rest_long, variable == "1_1_u_1")

cor(rest_11u1$diagnostic_judgment, rest_11u1$value)

cor.test(rest_11u1$diagnostic_judgment, rest_11u1$value)
```










