---
title: "Analysis Sample and Effect Size 11"
author: Simon Stephan
output: html_document
---

```{r setup, include=FALSE}
# packages
library(ez)
library(reshape2)
library(reshape)
library(ggplot2)
library(plyr)
library(pastecs)
library(data.table)
library(psych)
library(showtext)

font_add_google("Poppins", "Poppins")
font_add_google("Roboto Mono", "Roboto Mono")
showtext_auto()
```

```{r}
# read data 
#tdata.wide <- read.delim("dummy_data.txt", header=TRUE, sep="\t", na.strings="NA", dec=".", strip.white=TRUE)
tdata <- read.delim("Exp2_data.txt", header=TRUE, sep="\t", na.strings="NA", dec=".", strip.white=TRUE)
```

# Demographics
```{r}
mean(tdata$Age)
sd(tdata$Age)

colnames(tdata)[11] <- "Sex" # rename column listing subjects' sex
table(tdata$Sex) # 1 = male, 2 = female, 3 = non-binary
```

# Data preparation
```{r, message=FALSE}
# declare factors

tdata$Mechanisms <- factor(tdata$Mechanisms, levels = c("Independent", "Dependent"), labels = c("Non-intersecting", "Intersecting"))
tdata$Observation_Order <- factor(tdata$Observation_Order, levels = c("unobserved_first", "observed_first"))

tdata$diag_prob_rating <- (tdata$diag_prob_rating - 1) * 0.1

tdata_attributions <- subset(tdata, select = c(1:5))
tdata_diagnostics <- subset(tdata, select = c(1:3,6))

tdata_DependentAttributions <- subset(tdata, Mechanisms == "Intersecting")
tdata_DependentAttributions_given_correct_diags <- subset(tdata_DependentAttributions, diag_prob_rating == 1.0)
tdata_DependentAttributions_given_correct_diags <- subset(tdata_DependentAttributions_given_correct_diags, select = c(1:5))
tdata_DependentAttributions_given_Incorrect_diags <- subset(tdata_DependentAttributions, diag_prob_rating < 1.0)
tdata_DependentAttributions_given_Incorrect_diags <- subset(tdata_DependentAttributions_given_Incorrect_diags, select = c(1:5))




# rename columns 
colnames(tdata_attributions)
colnames(tdata_attributions)<- c("sID", "Mechanisms", "Observation_Order", "1_1_u_1", "1_u_u_1")

colnames(tdata_DependentAttributions_given_correct_diags)<- c("sID", "Mechanisms", "Observation_Order", "1_1_u_1", "1_u_u_1")
colnames(tdata_DependentAttributions_given_Incorrect_diags)<- c("sID", "Mechanisms", "Observation_Order", "1_1_u_1", "1_u_u_1")
```

```{r}
# built a new data set that has as a new factor coding whether subjects made correct vs. incorrect diagnostic judgments in the dependent-mechanisms condition

diagnostic_judgment <- c(rep("correct", 21))

tdata_DependentAttributions_given_correct_diags$diagnostic_judgment <- diagnostic_judgment

diagnostic_judgment <- c(rep("incorrect", 23))

tdata_DependentAttributions_given_Incorrect_diags$diagnostic_judgment <- diagnostic_judgment

tdata_DependentAttributions_givenDiag <- rbind(tdata_DependentAttributions_given_correct_diags, tdata_DependentAttributions_given_Incorrect_diags)

tdata_DependentAttributions_givenDiag_long <- melt(tdata_DependentAttributions_givenDiag, id=c("sID", "Mechanisms", "Observation_Order", "diagnostic_judgment"))

tdata_DependentAttributions_givenDiag_long$value <- (tdata_DependentAttributions_givenDiag_long$value - 1) * 0.1

tdata_DependentAttributions_givenDiag_long$variable <- factor(tdata_DependentAttributions_givenDiag_long$variable, levels = c("1_u_u_1", "1_1_u_1"))
```



```{r, message=FALSE}
tdata_attributions_long <- melt(tdata_attributions, id=c("sID", "Mechanisms", "Observation_Order"))

tdata_attributions_long$variable <- factor(tdata_attributions_long$variable, levels = c("1_u_u_1", "1_1_u_1"))

# 5. recode dependent variables (because Subj. answered on a 1 to 11 scale but we want 0 to 10)
tdata_attributions_long$value <- (tdata_attributions_long$value - 1) * 0.1

tdata_attributions_long["title"] <- c(rep("Human judgments", length(tdata_attributions_long$sID)))

data_wide <- dcast(tdata_attributions_long, sID + Mechanisms + Observation_Order + title ~ variable, value.var="value")

```

# Make bar charts 


## Chart showing the predictions 


```{r, echo = FALSE}
# create a theme to handle the design 
myTheme <- theme(plot.title = element_text(face="bold", size = 12),
        axis.title.x = element_text(face = "bold", size = 10),
        axis.title.y = element_text(face = "bold", size = 10),
        axis.text.x = element_text(size = 6, angle = 45, hjust = 1), 
        axis.text.y = element_text(size = 4, angle = 90),
        legend.text = element_text(size = 12),
        legend.title = element_text(face = "bold", size = 12),
        strip.text.x = element_text(size = 12),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line.x = element_line(colour = "black"), 
        axis.line.y = element_line(colour = "black"),
        axis.text = element_text(colour ="black"), 
        axis.ticks = element_line(colour ="black"))

```


## predictions based on integration of information from both potential causes

```{r}
# derive the model predictions for the "independent mechanisms" condition

##### Causal strength parameters

# base rates of root causes C and A
bc <- 0.7
ba <- bc

# strength between root causes and mechanism nodes
wcmc <- 0.6
wama <- wcmc

# strength of mechanism component
wmce <- 0.9 # the strong path
wmae <- wmce


alpha_symmetric <- 0.5 # for x_u_1_u_x




##### The 2 different test cases we need to model

#############################################
# legend: 
# t = telegraph (superior mech component)
# p = pony rider (inferior mech component)
# 1 = variable present 
# 0 = variable absent 
# u = variable unobserved
#############################################

#     C_Mc_E_Ma_A

#  1) 1_u_1_u_1
#  2) 1_1t_1_u_1


##################################################

# function for generalized power model of singular causation 

generalized_model <- function(target_power, alternative_power, alpha) 
{
  # formula 
  (target_power - target_power * alternative_power * alpha)/(target_power + alternative_power - target_power * alternative_power)
}

# 1) 1_u_1_u_1
power_target <- wcmc * wmce
power_alt <- wama * wmae
alpha <- alpha_symmetric

singular_1 <- generalized_model(power_target, power_alt, alpha)


# 1) 1_1_1_u_1
power_target <- wmce # simply the determined link strength
power_alt <- wama * wmae
alpha <-  alpha_symmetric


singular_2 <- generalized_model(power_target, power_alt, alpha)
```


```{r}
### predictions for the dependent-mechanisms condition

# we need to find out how likely it is that the target instead of the alternative cause activated the shared mechanism variable


# base rates of root causes C and A
bc <- 0.7
ba <- bc

# strength between root causes and mechanism nodes
wcmcma <- 0.6
wama <- wcmcma

# strength of mechanism components
wmcmae <- 0.9 # the strong path
wmamcma <- 0.7745967


alpha_symmetric <- 0.4 # for x_u_1_u_x


power_target <- wcmc #simply the determined link strength
power_alt <- wama * wmamcma
alpha <-  0.4



singular_3 <- generalized_model(power_target, power_alt, alpha)
singular_4 <- singular_3


```


```{r}
# create a data file containing the predictions 

Mechanisms <- c(rep("Non-intersecting",2), rep("Intersecting",2))
#Parameter_Info <- c(rep("no",4), rep("yes",4),rep("no",4), rep("yes",4))
#Observation_Order <- c(rep(c("unobserved_first", "unobserved_first", "observed_first", "observed_first"),4))

variable <- c(rep(c("1_u_u_1", "1_1_u_1", "1_1_u_1", "1_u_u_1"),1))

value <- c(singular_1, singular_2, singular_3, singular_4)

title <- c(rep("Ordinal predictions: integration of target and alternative cause", length(value)))

preds <- data.frame(Mechanisms, variable, value, title)

preds$Mechanisms <- factor(preds$Mechanisms, levels = c("Non-intersecting", "Intersecting"))
preds$variable <- factor(preds$variable, levels = c("1_u_u_1", "1_1_u_1"))

```



```{r}
theme_set(theme_bw(base_size = 14, base_family = "Poppins"))

predictions <- ggplot(preds, aes(x=variable, y=value, group = Mechanisms, colour = Mechanisms)) + 
  ggtitle("a) Model predictions")+
  coord_cartesian(ylim=c(0,1)) + 
  #coord_flip()+
  #scale_y_continuous(limits = c(0,1), breaks=seq(0, 1, 0.25), expand = c(0,0)) +
  #scale_x_discrete(labels=c("Direct: one link", "Iindirect: two links", "Indirect: five links")) +
  #stat_summary(fun.y = mean, geom = "bar", position = "dodge", colour = "black") +
  #stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0, position=position_dodge(0.1), size = 1) +
  stat_summary(fun.y=mean, geom="line", position=position_dodge(0.1), size = 1, alpha = 0.7)+
  stat_summary(fun.y=mean, colour="grey20", fill = c("#005083", "#005083", "#f0b64d", "#f0b64d"), geom="point", shape = 21, size = 3,
               position=position_dodge(0.1))+
  #geom_jitter(position = position_jitterdodge(jitter.height = 0.01, dodge.width = 0.90), alpha = 0.2, colour = "black") +
  #stat_summary(fun.y = median, geom = "point", shape = 23, size = 3, position = position_dodge(width = 0.90), colour = "black") +
  #stat_summary(fun.data = mean_cl_boot, geom = "errorbar", position = position_dodge(width = 0.90), width = 0.2) +
  #geom_hline(yintercept=0.9) + 
  #ggtitle("Means (95% bootstr. CIs)") +
  #myTheme +
  theme(
      legend.position = "top",
      legend.title = element_text(size = 14),
      legend.text = element_text(size = 12),
      axis.title = element_text(size = 13),
      axis.text.y = element_text(angle=90, hjust = 0.3, color = "black", size = 12),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      panel.grid = element_blank(),
      #panel.grid = element_line(colour="grey", linetype="dashed")
      )+
  #theme(axis.text.x = element_blank(), axis.title.x = element_blank())+
  labs(x = "Observed singular case", y = "Confidence that C = 1 caused E = 1") +
  scale_color_manual(values=c("#005083", "#f0b64d"))

predictions

#ggsave("predictions1.svg",width=5,height=4)
#ggsave("predictions1.pdf",width=5,height=4)
```


```{r}
# plot the interaction effect together with its 95% CI 

t <- qt(0.975, 86, lower.tail = TRUE, log.p = FALSE)
#t

effect <- "Delta D"
D_Mdiff <- (singular_2 - singular_1) - (singular_3 - singular_4)
SE <- 0.0
CI <- SE*t
CI_low <- D_Mdiff - CI
CI_up <- D_Mdiff + CI

DMdiff <- data.frame(effect, D_Mdiff, CI_low, CI_up)

ggplot(DMdiff, aes(x=effect, y=D_Mdiff)) +
    #coord_cartesian(ylim=c(-0.02,0.15)) +
    #scale_y_continuous(limits = c(0,0.2), breaks=seq(0, 0.2, 0.1), expand = c(0,0)) +
    #geom_line() +
    geom_errorbar(width=.0, aes(ymin=DMdiff$CI_low, ymax=DMdiff$CI_up)) +
    geom_point(shape=17, size=3, color = "black") +
    #ylim(-0.02,0.2)+
    scale_y_continuous(limits=c(-0.02, 0.4), breaks = seq(-0.02, 0.4, 0.08))+
    geom_hline(yintercept=0, linetype="dashed", color = "red")+
  theme(
      legend.position = "none",
      axis.line.y = element_line(colour = "black"),
      axis.title.x = element_blank(),
      axis.ticks.y = element_line(color = "black"),
      axis.ticks.x = element_blank(),
      #axis.ticks.y = element_line(colour = "black"),
      axis.text.x = element_text(size = 12, color = "black"),
      axis.title.y = element_blank(),
      axis.text.y = element_text(size = 14, color = "black"),
      panel.grid = element_blank(),
      panel.border = element_blank(),
      #panel.grid = element_line(colour="grey", linetype="dashed")
      )

#ggsave("interaction_pred.pdf",width=1.2,height=2)
#ggsave("interaction_pred.svg",width=1.2,height=2)
```



# Subjects' answers

## Descriptive Statistics
```{r}
library(pastecs)
by(tdata_attributions_long$value, list(tdata_attributions_long$Mechanisms, tdata_attributions_long$variable), stat.desc , basic = FALSE)

```




## Chart showing human singular causation judgments

how within-subject error bars are to be computed can be found here: http://www.cookbook-r.com/Graphs/Plotting_means_and_error_bars_(ggplot2)/ 

```{r}
ggplot(tdata_attributions_long, aes(x=variable, y=value, group = Mechanisms, colour= Mechanisms)) + 
  ggtitle("b) Subjects' singular causation judgments")+
  coord_cartesian(ylim=c(0,1)) +
  #coord_flip()+
  #scale_y_continuous(limits = c(0,1), breaks=seq(0, 1, 0.25), expand = c(0,0)) +
  #scale_x_discrete(labels=c("Direct: one link", "Iindirect: two links", "Indirect: five links")) +
  #stat_summary(fun.y = mean, geom = "bar", position = "dodge", colour = "black") +
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0, position=position_dodge(0.1), size = 1) +
  stat_summary(aes(y = value), fun.y=mean, geom="line", position=position_dodge(0.1), size = 1, alpha = 0.7)+
  stat_summary(aes(y = value), fun.y=mean, colour="grey20", fill = c("#005083", "#005083", "#f0b64d", "#f0b64d"), geom="point", shape = 21, size = 3, position=position_dodge(0.1))+
  #geom_jitter(alpha = 0.2, colour = "black") +
  #stat_summary(fun.y = median, geom = "point", shape = 23, size = 3, position = position_dodge(width = 0.90), colour = "black") +
  #geom_hline(yintercept=0.9) + 
  #ggtitle("Means (95% bootstr. CIs)") +
  theme(
      legend.position = "top",
      legend.title = element_text(size = 14),
      legend.text = element_text(size = 12),
      axis.title = element_text(size = 13, color = "white"),
      axis.text.y = element_text(angle=90, hjust = 0.3, color = "white", size = 12),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      panel.grid = element_blank(),
      #panel.grid = element_line(colour="grey", linetype="dashed")
      )+
  labs(x = "Observed singular case", y = "Confidence that C = 1 caused E = 1") +
  scale_color_manual(values=c("#005083",  "#f0b64d")) +
  theme(legend.position="top")



#ggsave("human_jugments.svg",width=5,height=4)
#ggsave("human_jugments.pdf",width=5,height=4)
```



```{r}
library(afex)
library(emmeans)

a1 <- aov_car(value ~ Mechanisms*variable + Error(sID/(variable)), tdata_attributions_long)
a1
```


```{r}
# same ANOVA as before
lmeModel <- lmer(value ~ Mechanisms*variable + (1|sID), data=tdata_attributions_long)

# follow-up analysis 

ls1 <- lsmeans(a1, c("variable", "Mechanisms")) # joint evaluation (basically gives the same table)
ls1
```


```{r}
############### 
# a conditional analysis 

ls2 <- lsmeans(a1, c("variable"), by = "Mechanisms") # group means by between-condition
ls2

# simple main effects 
pairs(ls2) # compares rep-measure differences separately for each between-factor level

# interaction contrast 
pairs(pairs(ls2), by = NULL)

#test(pairs(pairs(ls2), by = NULL), joint = TRUE) # This reproduces the F-Value of the ANOVA interaction


#lsmip(a1, High_Strength_Component ~ variable) # lsemans can also produce graphs
```

```{r}
# compute the confidence interval for the singular causation differences in each between-subject condition

# independent mechanisms condition

t <- qt(0.975, 86, lower.tail = TRUE, log.p = FALSE)
#t

effect <- "Mdiff"
Mdiff <- 0.357
SE <- 0.0433
CI <- SE*t
CI_low <- Mdiff - CI
CI_up <- Mdiff + CI

Mdiff
CI_low
CI_up


# dependent mechanisms condition

t <- qt(0.975, 86, lower.tail = TRUE, log.p = FALSE)
#t

effect <- "Mdiff"
Mdiff <- 0.168
SE <- 0.0433
CI <- SE*t
CI_low <- Mdiff - CI
CI_up <- Mdiff + CI

Mdiff
CI_low
CI_up

```



```{r}
# plot the interaction effect together with its 95% CI 

t <- qt(0.975, 86, lower.tail = TRUE, log.p = FALSE)
#t

effect <- "Delta D"
D_Mdiff <- 0.189
SE <- 0.0612
CI <- SE*t
CI_low <- D_Mdiff - CI
CI_up <- D_Mdiff + CI

DMdiff <- data.frame(effect, D_Mdiff, CI_low, CI_up)

ggplot(DMdiff, aes(x=effect, y=D_Mdiff)) +
    #coord_cartesian(ylim=c(-0.02,0.15)) +
    #scale_y_continuous(limits = c(0,0.2), breaks=seq(0, 0.2, 0.1), expand = c(0,0)) +
    #geom_line() +
    geom_errorbar(width=.0, aes(ymin=DMdiff$CI_low, ymax=DMdiff$CI_up)) +
    geom_point(shape=17, size=3, color = "black") +
    #ylim(-0.02,0.2)+
    scale_y_continuous(limits=c(-0.02, 0.4), breaks = seq(-0.02, 0.4, 0.08))+
    geom_hline(yintercept=0, linetype="dashed", color = "red")+
theme(
      legend.position = "none",
      axis.line.y = element_line(colour = "black"),
      axis.title.x = element_blank(),
      axis.ticks.y = element_line(color = "black"),
      axis.ticks.x = element_blank(),
      #axis.ticks.y = element_line(colour = "black"),
      axis.text.x = element_text(size = 12, color = "black"),
      axis.title.y = element_blank(),
      axis.text.y = element_text(size = 14, color = "black"),
      panel.grid = element_blank(),
      panel.border = element_blank(),
      #panel.grid = element_line(colour="grey", linetype="dashed")
      )

#ggsave("interaction.svg",width=1.2,height=2)
#ggsave("interaction.pdf",width=1.2,height=2)


```



## Analysis of diagnostic probability judgments 


## Chart showing human diagnostic judgments 

```{r}
diag_judgments_jit <- ggplot(tdata_diagnostics, aes(Mechanisms, diag_prob_rating, color = Mechanisms, fill = Mechanisms)) + 
  ggtitle("a) Subjects' diagnostic probability judgments") +
  #coord_cartesian(ylim=c(-0.1,2)) + 
  scale_y_continuous(limits = c(-0.05, 1.06), breaks=seq(0, 1, 0.25), expand = c(0,0)) +
  scale_x_discrete()+
  #annotate("rect", xmin=1.7, xmax=2.3, ymin=0.95, ymax=1.05, color="#31a354", fill = "white", size = 1) +
  #annotate("rect", xmin=1.7, xmax=2.3, ymin=0.45, ymax=0.55, color="#de2d26", fill = "white", size = 1) +
  geom_jitter(width = 0.3, height = 0.02, alpha = 0.6) +
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", position = position_dodge(width = 0.90), width = 0, size = 1) +
  stat_summary(fun.y=mean, colour="grey20", geom="point", shape = 21, size = 3)+
  #ggtitle("Means (95% bootstr. CIs)") +
  theme(
      legend.position = "none",
      axis.title = element_text(size = 12),
      #axis.text.x = element_blank(),
      #axis.title.x = element_blank(),
      panel.grid = element_blank(),
      #panel.grid = element_line(colour="grey", linetype="dashed")
      )+
  #theme(axis.text.x = element_text(size = 10, angle = 0, hjust = 0.5))+
  labs(x = "Mechanisms", y = "Rating: Diagnostic Probability of\n Target Mechanism given Effect") +
  scale_color_manual(values=c("#005083",  "#f0b64d")) +
  scale_fill_manual(values=c("#005083",  "#f0b64d")) +
  theme(legend.position="none")

diag_judgments_jit

#ggsave("diagnostic_judgments_jit.svg",width=5,height=3.8)
#ggsave("diagnostic_judgments_jit.pdf",width=5,height=3.8)
```

```{r}
t.test(tdata$diag_prob_rating ~ tdata$Mechanisms, alternative = "two.sided", var.equal = T, conf.level=0.95)

M_Diff <- 0.7272727 - 0.4886364 

CI <- c("0.1392229","0.3380498")
```




```{r}
# graph for dependent group with diag-judgments as control
library(scales) 

ggplot(tdata_DependentAttributions_givenDiag_long , aes(x=variable, y=value, group = diagnostic_judgment, colour= diagnostic_judgment)) + 
  ggtitle("b) grouped by quality of diagnostic probability rating") +
  coord_cartesian(ylim=c(0,1)) +
  #coord_flip()+
  #scale_y_continuous(limits = c(0,1), breaks=seq(0, 1, 0.25), expand = c(0,0)) +
  #scale_x_discrete(labels=c("Direct: one link", "Iindirect: two links", "Indirect: five links")) +
  #stat_summary(fun.y = mean, geom = "bar", position = "dodge", colour = "black") +
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0, position=position_dodge(0.1), size = 1) +
  stat_summary(aes(y = value), fun.y=mean, geom="line", position=position_dodge(0.1), size = 1)+
  stat_summary(aes(y = value), fun.y=mean, colour="grey20", fill = c("#ffd28f", "#ffd28f", "#c99f6e", "#c99f6e"), geom="point", shape = 21, size = 3, position=position_dodge(0.1))+
  #geom_jitter(alpha = 0.2, colour = "black") +
  #stat_summary(fun.y = median, geom = "point", shape = 23, size = 3, position = position_dodge(width = 0.90), colour = "black") +
  #geom_hline(yintercept=0.9) + 
  #ggtitle("Means (95% bootstr. CIs)") +
  #myTheme +
  labs(x = "Observed singular case", y = "Confidence that C = 1 caused E = 1") +
  scale_color_manual(name  ="Diagnostic judgment", labels=c("correct", "incorrect"), values=c("#ffd28f",  "#c99f6e")) +
  #scale_colour_discrete(name  ="Diagnostic judgment", labels=c("correct", "incorrect")) +
  theme(
      legend.position = "top",
      legend.title = element_text(size = 14),
      legend.text = element_text(size = 12),
      axis.title = element_text(size = 13, color = "black"),
      axis.text.y = element_text(angle=90, hjust = 0.3, color = "black", size = 12),
      axis.text.x = element_blank(),
      axis.title.x = element_blank(),
      panel.grid = element_blank(),
      #panel.grid = element_line(colour="grey", linetype="dashed")
      )


#ggsave("human_judgments_dependentCon.svg",width=5,height=4)
#ggsave("human_judgments_dependentCon.pdf",width=5,height=4)


```


```{r}
library(afex)
library(emmeans)

a1 <- aov_car(value ~ diagnostic_judgment*variable + Error(sID/(variable)), tdata_DependentAttributions_givenDiag_long)
a1
```


```{r}
# same ANOVA as before
lmeModel <- lmer(value ~ diagnostic_judgment*variable + (1|sID), data=tdata_DependentAttributions_givenDiag_long)

# follow-up analysis 

ls1 <- lsmeans(a1, c("variable", "diagnostic_judgment")) # joint evaluation (basically gives the same table)
ls1
```


```{r}
############### 
# a conditional analysis 

ls2 <- lsmeans(a1, c("variable"), by = "diagnostic_judgment") # group means by between-condition
ls2

# simple main effects 
pairs(ls2) # compares rep-measure differences separately for each between-factor level

# interaction contrast 
pairs(pairs(ls2), by = NULL)

#test(pairs(pairs(ls2), by = NULL), joint = TRUE) # This reproduces the F-Value of the ANOVA interaction


#lsmip(a1, High_Strength_Component ~ variable) # lsemans can also produce graphs
```


```{r}
# plot the interaction effect together with its 95% CI 

t <- qt(0.975, 42, lower.tail = TRUE, log.p = FALSE)
#t

effect <- "Delta D"
D_Mdiff <- 0.185
SE <- 0.0785
CI <- SE*t
CI_low <- D_Mdiff - CI
CI_up <- D_Mdiff + CI

DMdiff <- data.frame(effect, D_Mdiff, CI_low, CI_up)

ggplot(DMdiff, aes(x=effect, y=D_Mdiff)) +
    #coord_cartesian(ylim=c(-0.02,0.15)) +
    #scale_y_continuous(limits = c(0,0.2), breaks=seq(0, 0.2, 0.1), expand = c(0,0)) +
    #geom_line() +
    geom_errorbar(width=.0, aes(ymin=DMdiff$CI_low, ymax=DMdiff$CI_up)) +
    geom_point(shape=17, size=3, color = "black") +
    #ylim(-0.02,0.2)+
    scale_y_continuous(limits=c(-0.02, 0.4), breaks = seq(-0.02, 0.4, 0.08))+
    geom_hline(yintercept=0, linetype="dashed", color = "red")+
theme(
      legend.position = "none",
      axis.line.y = element_line(colour = "black"),
      axis.title.x = element_blank(),
      axis.ticks.y = element_line(color = "black"),
      axis.ticks.x = element_blank(),
      #axis.ticks.y = element_line(colour = "black"),
      axis.text.x = element_text(size = 12, color = "black"),
      axis.title.y = element_blank(),
      axis.text.y = element_text(size = 14, color = "black"),
      panel.grid = element_blank(),
      panel.border = element_blank(),
      #panel.grid = element_line(colour="grey", linetype="dashed")
      )

#ggsave("interaction_based_on_diag.svg",width=1.2,height=2)
#ggsave("interaction_based_on_diag.pdf",width=1.2,height=2)
```








